---
title: "Applied Data Analysis Lab Week 5"
output: html_notebook
---

# Exploring baseball data

```{r include = F}
require(tidyverse)
```

## Load dataset (Step 1&2)
```{r}
# batting <- Lahman::Batting
batting <- read_csv("adas/w5_prj/raw_data/batting.csv")
str(batting)
```

## Explore dataset (Step 3)
```{r}
head(batting)
# tail(batting)
```

## Filter by conditions (Step 4)
```{r}
batting_filtered <- batting %>%
    filter(between(yearID, 2008, 2018))

nrow(batting_filtered)

batting %>%
    filter(yearID == 2018)

batting %>%
    filter(yearID == 2018 | yearID == 2008)
```
## Summary (Step 5)
```{r}
batting_teams <- batting_filtered %>%
    group_by(yearID, teamID) %>%
    summarise_at(vars(c(AB:GIDP)), sum)
```
## Mutate (Step 6)
```{r}
batting_teams <- batting_teams %>%
    mutate(X1B = H - HR - X2B - X3B)
```
## Distribution (Step 7)
```{r}
batting_teams %>%
    ggplot(aes(x = R)) +
        geom_histogram(
            binwidth = 25,
            colour = "black",
            fill = "blue"
        )
```
## Distribution with density curve (Step 8)
```{r}
batting_teams %>%
    ggplot(aes(x = R)) +
        geom_histogram(
            mapping = aes(y = ..density..)
            binwidth = 25,
            colour = "black",
            fill = "blue"
        ) +
        geom_density(
            alpha = 0.3,
            fill = "magenta"
        )
```
## Scatter plot (Step 9)
```{r}
batting_teams %>%
    ggplot(aes(x = X1B, y = R)) +
        geom_point(
            colour = "darkgreen",
            shape = 4,
            size = 2
        )

batting_teams %>%
    ggplot(aes(x = X2B, y = R)) +
        geom_point(
            colour = "dodgerblue",
            shape = 1,
            size = 2
        )

batting_teams %>%
    ggplot(aes(x = X3B, y = R)) +
        geom_point(
            colour = "purple",
            shape = 5,
            size = 3
        )
```
## Scatter with colour gradient (Step 10)
```{r}
batting_teams %>%
    ggplot(aes(x = X1B, y = R)) +
        geom_point(aes(colour = AB, size = AB)) +
        scale_color_gradient(low = "#56B1F7", high = "#132B43")

batting_teams %>%
    ggplot(aes(x = X2B, y = R)) +
        geom_point(aes(colour = AB, size = AB)) +
        scale_color_gradient(low = "#56B1F7", high = "#132B43")

batting_teams %>%
    ggplot(aes(x = X3B, y = R)) +
        geom_point(aes(colour = AB, size = AB)) +
        scale_color_gradient(low = "#56B1F7", high = "#132B43")
```
## On-Base Percentage (Step 11)
```{r}
batting_teams <- batting_teams %>%
    mutate(OBP = (H + BB) / (AB + BB))
```
## OBP distribution (Step 12)
```{r}
batting_teams %>%
    ggplot(aes(x = OBP)) +
        geom_histogram(
            binwidth = 0.01,
            colour = "black",
            fill = "magenta"
        )
```
## OBP distribution with density curve (Step 13)
```{r}
batting_teams %>%
    ggplot(aes(x = OBP)) +
        geom_histogram(
            mapping = aes(y = ..density..),
            binwidth = 0.01,
            colour = "black",
            fill = "magenta"
        ) +
        geom_density(
            alpha = 0.3,
            fill = "dodgerblue"
        )
```
## `if_else()` function (Step 14)
```{r}
batting_teams <- batting_teams %>%
    mutate(
        OBP_grouped = if_else(
            OBP > mean(OBP),
            "above_average_OBP",
            "below_average_OBP"
        )
    )
```
## Boxplot (Step 15)
```{r}
batting_teams %>%
    ggplot(aes(x = OBP_grouped, y = R)) +
        geom_boxplot(aes(fill = OBP_grouped))
```
## Grouped distribution (Step 16)
```{r}
batting_teams %>%
    ggplot(aes(x = R, colour = OBP_grouped)) +
        geom_freqpoly()

batting_teams %>%
    ggplot(aes(x = R, fill = OBP_grouped)) +
        geom_histogram(colour = "black", binwidth = 20) +
        facet_wrap(~ OBP_grouped, nrow = 2) +
        theme(legend.position = "none")
```
## Linear regression (Step 17)
```{r}
batting_teams %>%
    ggplot(aes(x = OBP, y = R)) +
        geom_point(colour = "dodgerblue", size = 2) +
        geom_smooth(method = "lm", colour = "magenta")
```
