---
title: "Applied Data Analysis in Sport"
subtitle: " Assessment 4: Reproducible Data Analysis Project"
author: "Data Man"
date: "Semester 1 2022"
output:
  pdf_document:
    theme: readable
    highlight: breezedark
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE)
```
```{r}
require(tidyverse)
require(ggplot2)
```


# Introduction
## a) background
## b) case description
## c) project target
## d) justification and importance

# Reading and cleaning the raw data
```{r}
# Unzip the raw data file.
file_dir <- "data/project-data/"

if (!dir.exists(file_dir)) {
  unzip("data/raw/project-data.zip", exdir = "data")
}

# Load the raw datasets.
filenames <- list.files(file_dir)

for (name in filenames) {
  assign(
    str_extract(name, "(?<=a_).+(?=\\.)"),
    read_csv(paste0(file_dir, name))
  )
}
```
```{r}
str(df_salaries)
```
```{r}
str(df_player_stat)
```
```{r}
str(df_team_stat_1)
```
```{r}
str(df_team_stat_2)
```
```{r}
str(df_payroll)
```


# Exploratory analysis
## a) checking for errors and missings
## b) checking for variable distributions
## c) checking for variable relationships
## d) justification about modelling

# Data modelling and results
## a) modelling
## b) assumption checking
## c) output and interpretation

# Player recommendations

# Summary

# Reference