---
title: "Introduction to Data Science G (11516)"
subtitle: "Final Assesment"
author: ""
date: "Semester 1, 2021"
output:
  pdf_document:
    highlight: breezedark
    df_print: kable
---

```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```
# Part B - Data Preparation, exploring and modeling

```{r include=FALSE}
require(tidyverse)
require(ggplot2)
require(GGally)
library(caret)

#' ## Task 1: Data Preparation and Wrangling

#' ### 1
covid19 <- read_csv("Covid19.csv")
tests <- read_csv("Tests.csv")
countries <- read_csv("Countries.csv")
recovered <- read_csv("Recovered.csv")

#' ### 2
recovered <- gather(recovered, key, value, 2:106)

#' ### 3
colnames(covid19) <- c("Code", "Country", "Continent",
                       "Date", "NewCases", "NewDeaths")

colnames(tests) <- c("Code", "Date", "NewTests")

colnames(countries) <- c("Code", "Country", "Population", "GDP", "GDPCapita")

colnames(recovered) <- c("Country", "Date", "Recovered")

#' ### 4
covid19[["Date"]] <- as.Date(covid19[["Date"]], format = "%Y-%m-%d")

tests[["Date"]] <- as.Date(tests[["Date"]], format = "%Y-%m-%d")

recovered[["Date"]] <- recovered[["Date"]] %>%
    str_replace_all("\\.", "-") %>%
    as.Date(format = "%Y-%m-%d")

#' ### 5
covid19 <- tests %>%
    merge(covid19, all = T) %>%
    merge(recovered, all = T) %>%
    merge(countries, by = c("Code", "Country"), all = T)

#' ### 6
colSums(is.na(covid19))

colSums(is.na(tests))

colSums(is.na(countries))

colSums(is.na(recovered))

covid19[is.na(covid19)] <- 0

#' ### 7
covid19 <- covid19 %>%
    mutate(Week = as.numeric(format(Date, "%W"))) %>%
    mutate(Month = as.numeric(format(Date, "%m")))
```

## Task 2: Exploratory Data Analysis
```{r include=FALSE}
### 1
covid19 <- covid19 %>%
    arrange(Date, Country) %>%
    group_by(Country) %>%
    mutate(CumCases = cumsum(NewCases),
           CumDeaths = cumsum(NewDeaths),
           CumRecovered = cumsum(Recovered),
           CumTests = cumsum(NewTests)) %>%
    ungroup()

### 2
covid19 <- covid19 %>%
    mutate(Active = CumCases - (CumDeaths + CumRecovered),
           FatalityRate = CumDeaths / CumCases)

### 3
covid19 <- covid19 %>%
    mutate(Cases_1M_Pop =  CumCases * (10^6) / Population,
           Deaths_1M_Pop = CumDeaths * (10^6) / Population,
           Recovered_1M_Pop = CumRecovered * (10^6) / Population,
           Tests_1M_Pop = CumTests * (10^6) / Population)

```

## 4.
The day with highest reported death toll:
```{r}
covid19 %>%
    group_by(Date) %>%
    summarize(DeathToll = sum(NewDeaths)) %>%
    filter(DeathToll == max(DeathToll))
```

## 5.
Here is the graph of how the cumulative data change over the time for the whole world:

```{r}
covid19 %>%
    group_by(Date) %>%
    summarize(Case = sum(NewCases),
              Death = sum(NewDeaths),
              Recovered = sum(Recovered),
              Test = sum(NewTests)) %>%
    gather("Column", "Count", "Case":"Test") %>%
    ggplot(aes(Date, log(Count), group = Column, col = Column)) +
        geom_line() +
        ggtitle("COVID-19 Cases (2020)") +
        xlab("Date") +
        ylab("Count (log)") +
        scale_colour_viridis_d(name = "") +
        theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

```{r include=FALSE}
#' ### 6
lastDay_data <- covid19 %>% filter(Date == "2020-05-05")

#' ### 7
# extract whole records of the top 10 countries with highest active cases
top10active <- lastDay_data %>%
    arrange(desc(Active)) %>%
    head(10) %>%
    select(Country)

top10activeW <- covid19 %>%
    filter(Country %in% top10active$Country)

# extract whole records of the top 10 countries with highest confirmed cases
top10case <- lastDay_data %>%
    arrange(desc(CumCases)) %>%
    head(10) %>%
    select(Country)

top10caseW <- covid19 %>%
    filter(Country %in% top10case$Country)

# extract whole records of the top 10 countries with highest fatality rate
top10fatality <- lastDay_data %>%
    arrange(desc(FatalityRate)) %>%
    head(10) %>%
    select(Country)

top10fatalityW <- covid19 %>%
    filter(Country %in% top10fatality$Country)

# extract whole records of the top 10 countries with highest test number
top10test <- lastDay_data %>%
    arrange(desc(Tests_1M_Pop)) %>%
    head(10) %>%
    select(Country)

top10testMW <- covid19 %>%
    filter(Country %in% top10test$Country)
```

## 8.
Confirmed, death, recovered cases and test number for each continent:
```{r}
continent_data <- lastDay_data %>%
    group_by(Continent) %>%
    filter(Continent != 0) %>%
    summarize(Confirmed = sum(CumCases),
              Death = sum(CumDeaths),
              Recovered = sum(CumRecovered),
              Test = sum(CumTests))

continent_data
```

## 9.
Total number of cases over the time for the top 10 countries:

```{r}
top10caseW %>%
    ggplot(aes(Date, log(CumCases), group = Country, col = Country)) +
        geom_line() +
        ggtitle("COVID-19 Total Cases (2020)") +
        xlab("Date") +
        ylab("Count (log)") +
        scale_colour_viridis_d() +
        theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

## 10.
New cases, new deaths and new recovered cases over time for each country of top 10 highest current active cases:

```{r}
top10activeW %>%
    gather("column", "value", "NewCases":"Recovered") %>%
    ggplot(aes(Date, log(value), group = column, col = column)) +
        geom_line() +
        ggtitle("COVID-19 Cases (2020)") +
        xlab("Date") +
        ylab("Count (log)") +
        scale_colour_viridis_d(name = "",
                               label = c("confirmed", "death", "recovered")) +
        theme(plot.title = element_text(hjust = 0.5, face = "bold"),
              legend.position = "bottom") +
        facet_wrap(. ~ Country, ncol = 3)
```

## 11.
Total number of infected cases, total tests and total tests per million for the top 10 countries with highest total tests per one million population:

```{r}
top10testMW %>%
    filter(Date == "2020-05-05") %>%
    gather("column", "value",
               c("CumCases", "CumTests", "Tests_1M_Pop")) %>%
    ggplot(aes(column, value, fill = column)) +
        geom_col() +
        ggtitle("COVID-19 Cases (2020)") +
        xlab("Total Number") +
        ylab("Count") +
        scale_x_discrete(labels = c("infacted", "test", "test/mil")) +
        scale_fill_viridis_d(name = "Total Number",
                             label = c("infacted", "test", "test/mil")) +
        theme(plot.title = element_text(hjust = 0.5, face = "bold"),
              legend.position = "bottom") +
        facet_wrap(. ~ Country, ncol = 3)
```

## 12.
Statistics of all continents:

```{r}
continent_data %>%
    gather("column", "value", -Continent) %>%
    ggplot(aes(Continent, log(value), fill = column)) +
        geom_col(position = "dodge") +
        ggtitle("COVID-19 Cases (2020)") +
        xlab("Continent") +
        ylab("Count (Log)") +
        scale_fill_viridis_d(name = "") +
        theme(plot.title = element_text(hjust = 0.5, face = "bold"),
              axis.text.x = element_text(angle = 45, vjust = 0.5))
```

## Task 3: Data-Driven Modeling
```{r}
#' ### 1
cor_data <- lastDay_data %>%
    select(CumCases, CumTests, Population, GDP, GDPCapita)
```

## 2.
We can see that total cases have very high correlation coefficient with total test numbers and GDP:

```{r}
ggcorr(cor_data, label = T)
```

## 3.
We can see that the distribution of cumulative cases (`CumCases`) does not follow normal distribution:

```{r}
cor_data %>%
    ggplot(aes(CumCases)) +
        geom_histogram(bins = 50) +
        ggtitle("Distribution of Cumulative Cases") +
        xlab("Cumulative Cases") +
        theme(plot.title = element_text(hjust = 0.5, face = "bold"))

cor_data %>%
    ggplot(aes(log(CumCases))) +
        geom_histogram(aes(y = ..density..), bins = 20) +
        geom_density(color = "red") +
        ggtitle("Distribution of Cumulative Cases (Log)") +
        xlab("Cumulative Cases (Log)") +
        theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```

## 4.
Here are the outliers of the variable cumulative cases `CumCases`:

```{r}
boxplot.stats(cor_data$CumCases)$out
```

```{r}
#' ### 5
index <- createDataPartition(seq(nrow(cor_data)), p = .65, list = F)

train <- cor_data[index, ]
test <- cor_data[-index, ]
```

## 6.
```{r}
model_1 <- lm(CumCases ~ GDP, data = train)
summary(model_1)
```
We can see the R^2 is `r summary(model_1)$r.squared`, which means the explanatory variable `GDP` can explain around `r summary(model_1)$r.squared * 100`% of the variation in the response variable `CumSums`.

```{r}
plot(model_1)

ggplot(model_1, aes(model_1$residuals)) +
    geom_histogram(aes(y = ..density..), fill = "orange") +
    geom_density(color = "blue")
```
From the plots we can see that the model does not fit well, the distribution of residuals did not follow normal distribution.


Root mean square error value (`RMSE`):
```{r}
test$predict <- predict(model_1, test)

RMSE(test$CumCases, test$predict)
```

## 7.
```{r}
model_2 <- lm(CumCases ~ ., data = train)
summary(model_2)
```
We can see the R^2 is `r summary(model_2)$r.squared`, which means using all the other ones as explanatory variables can explain around `r summary(model_2)$r.squared * 100`% of the variation in the response variable `CumSums`.

```{r}
plot(model_2)

ggplot(model_2, aes(model_2$residuals)) +
    geom_histogram(aes(y = ..density..), fill = "orange") +
    geom_density(color = "blue")
```
From the plots we can see that the model does not fit well, the distribution of residuals did not follow normal distribution.


Root mean square error value (`RMSE`):
```{r}
test$predict <- predict(model_2, test)

RMSE(test$CumCases, test$predict)
```
The `RMSE` is better than previous model. 
